---
title: 'STATS 415: Homework 5'
author: "Kimberly A. Brink"
date: "February 24, 2016"
output: pdf_document
header-includes:
- \usepackage{fancyhdr}
- \usepackage{bm}
- \pagestyle{fancy}
- \fancyhead[RO,RE]{Kimberly A. Brink}
---

1. The textbook describes that the cv.glm() function can be used in order to compute the LOOCV test error estimate. Alternatively, one could compute those quantities using just the glm() and predict.glm() functions, and a for loop. You will now take this approach in order to compute the LOOCV error for a simple logistic regression model on the Weekly data set.

(a) Fit a logistic regression model that predicts Direction using
Lag1 and Lag2.

```{r, include=FALSE}
library(ISLR)
```

```{r 1a}
logit.fit <- glm(Direction ~ Lag1 + Lag2, data = Smarket, family = 'binomial')
summary(logit.fit)
```

**Neither Lag1 nor Lag2 significantly predicted direction ($p > 0.05$).**

(b) Fit a logistic regression model that predicts Direction using
Lag1 and Lag2 using all but the first observation.

```{r 1b}
head(Smarket[-1,])
logit.fit <- glm(Direction ~ Lag1 + Lag2, data = Smarket[-1,], family = 'binomial')
summary(logit.fit)
```

**I successfully removed the first observation and performed a logistic regression. Unsurprisingly, the outcome of the logistic regression was almost identical to the results in question 1.**

(c) Use the model from (b) to predict the direction of the first observation. You can do this by predicting that the first observation will go up if Pr(Direction=“Up” | Lag1, Lag2) > 0.5. Was this observation correctly classified?

```{r 1c}
contrasts(Smarket$Direction)
prediction = ifelse(predict(logit.fit, newdata = Smarket[1,], type="response")>0.5, "Up", "Down")
prediction == Smarket[1,]$Direction
```

**The observation was correctly classified.**

(d) Write a for loop from i = 1 to i = n, where n is the number of
observations in the data set, that performs each of the following
steps:

> i. Fit a logistic regression model using all but the ith observation to predict Direction using Lag1 and Lag2.

> ii. Compute the posterior probability of the market moving up
for the ith observation.

> iii. Use the posterior probability for the ith observation in order
to predict whether or not the market moves up.

> iv. Determine whether or not an error was made in predicting
the direction for the ith observation. If an error was made,
then indicate this as a 1, and otherwise indicate it as a 0.

```{r 1d}
for (i in 1:length(Smarket$Direction)){
  logit.fit <- glm(Direction ~ Lag1 + Lag2, data = Smarket[-i,], family = 'binomial')
  post.prob = predict(logit.fit, newdata = Smarket[i,], type="response")
  Smarket$prediction[i] = ifelse(post.prob>0.5, "Up", "Down")
}

Smarket$error = ifelse(Smarket$prediction != Smarket$Direction, 1, 0)
```

(e) Take the average of the n numbers obtained in (d)iv in order to
obtain the LOOCV estimate for the test error. Comment on the
results.

```{r 1e}
mean(Smarket$error)
```

**LOOCV demonstrates that, using this logistic regression, you are likely to make an error a little less than half the time (48.5%). This is not much better than randomly guessing. This is expected considering that neither Lag1 nor Lag2 were strong predictors of the direction of the stock market according to our results in question 1.**


